---
title: "seaaroundus"
author: "Leonie Grau, Felix Rentschler, Victoria Kolodziej"
date: "9 März 2018"
output:
  html_document:
    toc: yes
    toc_float: true
  pdf_document:
    number_sections: yes
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(shiny)
library(ggplot2)
library(tidyverse)
```

# Introduction

The Sea Around Us initiative is a research project by the University of British Columbia, that provides reported an reconstructed catch data for worldwide fisheries and fishery resources. Since overfishing is the major driver on the declining viability of fish stocks, a monitoring program with valid data is crucial for evaluating the marine development. The database thus provides information for policy makers, fishery managers and the scientific community to get a better understanding of global fish catches and more detailed information on which countries are involved. The impact of fishery on the marine ecosystems therefore is presented on a spatial scale in such a way that it has an ecological and political relevance. <cite>(Pauly D. and Zeller D. , 2015. Sea Around Us Concepts, Design and Data<cite> [www.seaaroundus.org](www.seaaroundus.org))

The database covers catch data over 64 years, from 1950 to 2014, classified by various categories such as taxon, fishing sector (industrial, artisanal, subsistence and recreational), catch type (landed and discarded catch), reporting status (reported and unreported) and so forth. The data can be shown for different spatial divisions such as Exclusive-economic zones (EEZ) and large marine ecosystems (LME) or organisational divisions such as regional fisheries management organizations (RFMO) or fishing-entity. Also, the data can be returned either in tonnage (tons) ore monetary value (US$). 
The data can be queried and looked at directly on the website www.seaaroundus.org where it can also be downloaded. For the purpose of a more in depth data work the R library “seaaroundus” is provided containing some functions to accesss the SAU-database. 

The aim of this project is to show worldwide catch data differentiated by country, i.e. the amount of catch each country has taken out of the oceans over time as well as differentiated by exclusive economic zones as a spatial indicator, i.e. how much fish is taken out where in the world. In addition, we want to show the proportion of discards in total catch in order to get an impression of the share of "waste" in worldwide fish takeouts. Moreover, this information shall be displayed for different time periods for the purpose of comparison and a deeper understanding of fish catch development over the past decades. 
In order to get the relevant data from the database we used the provided "seaaroundus"-package in R. For the visualisation of the data we used R Shiny and created an interactive (web) application.


# Materials and Methods

## Data extraction
The seaaroundus-package in R provides 6 functions to access the seaaroundus database and get different information. The main function to get specific catch data is the `catchdata`-function which transfers the input arguments to the database and returns either a data frame or a plot: 
```{r, eval=FALSE}
seaaroundus::catchdata(region, id, measure = "tonnage", dimension = "taxon", limit = 10, chart = FALSE, ...)
```

Input arguments specify the spatial dimension (`region`, `id`) as well as the type of data (`measure`, `dimension`), i.e. how it is quantified and in what categories it is presented. 
Options for `region` are 

* eez (exclusive economic zones)
* lme (lareg marine ecosystems)
* rfmo (regional fisheries management organisations)
* highseas
* fao
* eez-bordering
* fhishing entity
* taxon

The `id` depends on the specified region. A list of ids for each region can be called via the function `listregion(region)`.

Options for `dimension` are

* taxon
* commercialgroup
* functionalgroup
* country
* sector
* catchtype
* reporting-status
* layer

The arguments `limit` and `chart` give the possibilty for some queries to limit the returned data and to return it as a chart instead of a `data.frame`.

The `catchdata`-function does not provide the option to get information for more than one `id` at once. That's why a loop was run on the function passing a vector with all required ids in order to get overall information stored in one `data.frame`. This was done to get information on 

* the total catch for each country (`region="fishing-entity"` and `dimension="country"`)
* the amount of landings and discards for each country (`region="fishing-entity"` and `dimension="catchtype"`)
* the total catch *or* the share of discards within all Exclusive Economic Zones (EEZ) (`region="eez"` and `dimension="catchtype"`)

For all queries `measure` was set to `"tonnage"`. The queried data was put into `data.frames` and saved locally serving as database for the Shiny Web App. Below an exemplary code is shown. The whole code can be found in Appendix A. 

```{r, eval=F, echo=T}
#------------------------------
# Dataframe df_total_catch: Total fish catch in tons for every country and every year
#------------------------------
df_total_catch<- catchdata('fishing-entity', countryids[1], measure='tonnage', dimension='country')

for (i in 2:length(countryids)){ 
  #beginning with 2.column (first one is year)
  #every new column is the time series of total fish catches for one country:
  newcol<- catchdata('fishing-entity', countryids[i], measure='tonnage', dimension='country')
  df_total_catch[countrynames[i]] <-newcol[,2]
}
```

The shapefile of the worldwide EEZs was provided by the Marine Regions Geodatabase ([www.marineregions.org](www.marineregions.org)), which is managed by the Flanders Marine Institute. The map in the shiny app uses the version of 2014. Because neither the ids nor the country names correspond completely with the names of the SeaAroundUs data, the latter were adjusted manually. After this correction, the EEZ id of the SAU data could be added to the shapefile and be used as a key for merging. The SAU database contains 280 EEZs, while the Marine Regions database has only 249 EEZs. Because some zones within one database are split up into different zones in the other database or are missing, the zones shown in the map are reduces to 205 after the merge. 

## Visualisation

In order to visualise the queried data _R-Shiny Web App_ was used. This allows an interactive display of the information gathered on global fish catch. In general, the data is shown in a graph as well as in a table to get an overview of the numbers. To get an impression of the spatial distribution a map is displayed, too. All three outputs react to different dynamic input variables which the user is able to modify according to his or her interests. 

The implemented shiny app command consists of two arguments which need to be defined: 
`shinyApp(ui, server)` 
In the first one, the user interface is specified. Here, the layout contains a `sidebarPanel` and a `MainPanel` where input and ouput variables are arranged. The input variables build the interactive part providing a set of choices towards the ouput: three `select`-options and one `slider`-option give the possibility to change 

* the information which shall be displayed (*total catch* or _proportion of discards_)
* the time span for which the information shall be displayed (ranging from 1950 to 2014)
* the number of countries for which the information shall be displayed in the plots (ranging from 1 to 12)
* the number of entries to be listed in the tables (5 choices given)

The `MainPanel` defines the arrangement of the output in Tabs, i.e. showing either the plot *or* the table *or* the map. 
In the second argument of the shiny app the _server_ is defined, being the part which reacts to the varying input variables and which compiles the output. First, two reactive functions, `dataInput` and `calcTable`, return the changing data which are needed to produce the outputs. Then, the outputs (plot, table and map) are created. Therefore, the returned changed input data from `dataInput()` is further processed for the `output$tabGraph` via an `if else` statement and plotting commands using `ggplot()`. The return of the `calcTable`-function gets passed to the `output$tabTable` to create the respective table. In a last step, `output$tabMap`, the map is created via `leaflet` which allows the user to zoom in and out and have a closer look at the different eezs. The map is produced each time `dataInput$data_map` changes. 
The Shiny App is run via the command `shinyApp(ui = ui, server = server)`. 
The whole code for the Shiny App can be found in Appendix B.

# Results

The seaaroundus-Shiny Web App-R document is attached to the project file "seaaroundus" and can be opened and run via RStudio. In the following, exemplary screenshots are used to describe the outcome. 

The "total catch"-plot shows the development of the total amount of catch per year in tonnage over time as a stacked chart for up to 12 countries which are ordered descendingly by averaged amount of catch. For each chosen number of countries the remaining countries are summarised as "Others", accordingly. The default settings display 6 countries for the whole time period from 1950 to 2014, showing Japan to be the country with the highest average catch during this time span followed by the Russian Federation, Peru, USA, China and Norway (Figure 1). The table output provides the average catch per year and offers to return the data for 5, 10, 15, 20 or all countries, which in total is 197. Here, the default is set to 5 countries. 

------------------ INSERT FIGURE 1 HERE ----------------------

The "discards"-plot shows the development of the share of discards in total catch in percent over time as a line plot. Again, up to 12 countries can be displayed chosen by the rank they occupy when descendingly ordered by average share of discards. Kuwait has the highest amount of discards within the default time span followed by SaintPierre & Miquelon (France), El Salvador, Costa Rica, Portugal and Nicaragua which can be derived from the table (Figure 2).

------------------ INSERT FIGURE 2 HERE ----------------------

The eez-maps display the distribution of worldwide catches within 249 EEZs. The map with default settings for the *total_catch* data shows high catch rates around ... and .... whereas in ... and ... lower amounts of fish is taken out (Figure 3). The *discards* map shows lower rates of discard in ... and high shares of discards around .... (Figure 4).

------------------ INSERT FIGURE 3 HERE ----------------------
------------------ INSERT FIGURE 4 HERE ----------------------

# Discussion and Conclusion

The provided `catchdata`-function in the seaaroundus-package serves well to get differently combined information on global fish catch data. However, there are some limits to the function. For example, the `id` input argument only allows to put in one id leading to strong restrictions concerning ad-hoc information on more than one id within the specified `region` argument. Furthermore, some quite interesting combinations, though listed as possible, do not return any data at all, such as `catchdata(region="taxon", id="", measure="tonnage", dimension="country")`. For the purpose of a more free handling of the fish catch data and faster processing of required queries, a direct access to the database would have been desirable. However, within the scope of this project, it was not possible to easily work around the provided `catchdata`-function and create, e.g., a direct server connection. Hence, the desired information on catch data for all countries had to be performed via a loop processing huge amount of data locally thus taking up a relatively large amount of time. 

The visualisation of the queried data via R-Shiny provides interactive features allowing the viewer to define the output, i.e. adjust the data according to one's interests. Within the timeframe of this project, only a glimpse into the manifold possibilites of interactive visualisation provided by the shiny web app could be tested and implemented. There is, most certainly, a more efficient way to process the changing data input and transform it into the wanted output. Moreover, some further possibilities of outputs could be implemented ranging from additional ways to display the current data to adding more choices towards fish catch data.

To conclude, the seaaroundus database provides comprehensive data on global fish catch serving to answer different ecological, political and economical questions related to worldwide fishing. The data can be queried and downloaded from the website or accessed via the provided R-package "seaaroundus". However, both ways have limitations concerning some data queries. For the purpose of this course, a direct access to the database would have been preferable, making it possible to formulate personalised SQL-queries and shift computing effort which allows for more data processing.
However, the result of this project has to be considered against the background of a limited timeframe where several different materials and methods were taken into account. In any case, the personal perception concerning gain of knowledge and skills for each group member is: thumbs up.

# References

R Core Team (2017). R: A language and environment for statistical computing. R Foundation for Statistical Computing, Vienna, Austria. URL https://www.R-project.org/.

Scott Chamberlain and Robert Scott Reis (2017). seaaroundus: Sea Around Us API Wrapper. R package version 1.2.0. https://CRAN.R-project.org/package=seaaroundus

Winston Chang, Joe Cheng, JJ Allaire, Yihui Xie and Jonathan McPherson (2017). shiny: Web Application Framework for R. R package version 1.0.5. https://CRAN.R-project.org/package=shiny
