---
title: "Sea Around Us Project"
author: "Leonie Grau, Felix Rentschler, Victoria Kolodziej"
date: "March 6, 2018"
output: html_document
runtime: shiny
---

# Introduction

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(shiny)
library(ggplot2)
library(tidyverse)
```
The Sea Around Us initiative is a research project by the University of British Columbia, that offers reconstructed catch data for worldwide fisheries and fishery resources. Since overfishing is the major driver on the declining viability of fish stocks, a monitoring program with valid data is crucial for evaluating the marine development. The database provides information for policy makers, fishery managers and the scientific community to get a better understanding of global fishery catches and more detailed information about which countries are involved.  The impact of fishery on the marine ecosystems therefore is presented on a spatial scale in such a way that it has an ecological and political relevance. <cite>(Pauly D. and Zeller D. , 2015. Sea Around Us Concepts, Design and Data<cite> [www.seaaroundus.org](www.seaaroundus.org))

The database covers catch data over 64 years (1950-2014), 4 fishing sectors (industrial, artisanal, subsistence and recreational), 2 catch types (landed and discarded catch) and 2 types of reporting status (reported and unreported). There are 5 regions that can be shown: Exclusive-economic zones (EEZ), large marine ecosystems (LME), regional fisheries management organizations (RFMO),  EEZ-bordering and taxon. [...]

For the purpose of a more in depth data work the R library “seaaroundus” is provided. The aim of this project is to create an interactive web app to present specific data from the sea around us database. This is realised by the use of  the R package “shiny”. 


## Material and Methods

The basis for this project is the seaaroundus-R-package providing funtions to access the seaaroundus database. The main function is the `catchdata`-function which returns either a data frame or, for some queries, a plot.

```{r}
seaaroundus::catchdata
```

This function accessses the database via another function `callapi` tranferring the input arguments as a query and collects the 

## Results


```{r tabsets, echo=FALSE}

#read in data: ----
df_fishing_all <- source("df_fishing_all.Rdmpd")
df_fishing_all <- df_fishing_all[[1]]
df_Q2 <- read.delim("df_Q2")

# Define UI for seaaroundus app ----
ui <- fluidPage(
  
  # App title ----
  titlePanel("seaaroundus"),
  
  # Sidebar layout with input and output definitions ----
  sidebarLayout(
    
    # Sidebar panel for inputs ----
    sidebarPanel(
      
      # Input: Select-option for type of display in plots ----
      selectInput(inputId = "dim",
                  label = "Select type",
                  choices = c("total catch", "discards")),
      
      # Input: Slider for range of years to be displayed ----
      sliderInput(inputId = "range",
                  label = "Select time span",
                  min = 1950,
                  max = 2014,
                  sep = "",
                  value = c(1950, 2014)),
      
      # Input: Select-option for number of countries to be displayed in plots
      selectInput(inputId = "number",
                  label = "Select number of countries in Graph",
                  choices = c(1:12),
                  selected = 6),
      
      # Input: Select-option for number of entries to be displayed in tables
      radioButtons(inputId = "table_len",
                   label = "Select number of entries in Table",
                   inline=T,
                   choiceNames = c("5", "10", "15", "20", "all"),
                   choiceValues = c(5, 10, 15, 20, 197))
    ),
    
    # Main panel for displaying outputs with options ----
    mainPanel(
      
      # Output: Two tab panels: ----
      tabsetPanel(type="tabs",
                  tabPanel("Graph", plotOutput(outputId = "areaPlot")),
                  tabPanel("Table", tableOutput("values"))
      )
    )
  )
)


# Define server logic required to draw plots and show tables ----
server <- function(input, output) {
  
  # Reactive expression to create data frame of all input values ----
  # sliderValues <- reactive({
  #   
  #   data.frame(
  #     Name = c("Range"),
  #     Value = as.character(c(paste(input$range, collapse = " ")
  #     )),
  #     stringsAsFactors = FALSE)
  #   
  # })
  
  # First reactive function returning basis for plots and tables (output 1 and 2)
  dataInput <- reactive({
    
    data <- switch(input$dim,
                   "total catch" = df_fishing_all,
                   "discards" = df_Q2)
    
    range <- input$range
    
    number <- input$number
    
    data_plot <- data %>% filter(years>=range[1], years<=range[2])
    
    if(input$dim == "total catch"){
      data_table <- data_plot %>% gather(., key="country", value="tonnage", -c(years)) %>%
        group_by(country) %>% summarise(avg=sum(tonnage))
    }
    else {
      data_table <- data %>%
        mutate(perc_disc = (discards/(landings+discards))*100) %>% 
        group_by(country) %>% summarise(avg=mean(perc_disc))
    }
    
    return(list("data_plot"=data_plot, "data_table" = data_table, "number"=number))
  })
  
  # Second reactive expression returning values for tables (output 2) ----
  calcTable <- reactive({
    
    data_table <- dataInput()$data_table
    
    if(input$dim=="total catch"){
      data_table <- arrange(data_table, desc(avg))[c(1:input$table_len),]
      colnames(data_table) <- c("country", "sum of catches in tons")
    }
    else{
      data_table <- arrange(data_table, desc(avg))[c(1:input$table_len),]
      colnames(data_table) <- c("country", "average discards in %")
    }
    
    return(data_table)
  })
  
  
  # Output 1: Plots ----
  output$areaPlot <- renderPlot({
    
    data <- dataInput()$data_plot
    number <- as.integer(dataInput()$number)
    
    if (input$dim == "total catch"){
      
      data <- data %>% gather(., key="country", value="tonnage", -c(years)) 
      
      data_arranged <- data %>% group_by(country) %>% summarise(avg=mean(tonnage)) %>% arrange(., desc(avg))
      
      data_high <- data %>% filter(country %in% data_arranged[c(1:number),]$country)
      data_low <- data %>% filter(country %in% data_arranged[c((number +1):nrow(data)),]$country)  %>% group_by(years) %>%
        summarise(tonnage = sum(tonnage)) %>% mutate(country = "Others") %>%
        select(years,country,tonnage)
      
      
      data <- bind_rows(data_high, data_low)
      
      ggplot(data=data, aes(x= years, y=tonnage))+
        geom_area(aes(fill=factor(country, levels=c(data_arranged[c(1:number),]$country, "Others"))))+
        theme(legend.position = "right")+
        guides(fill=guide_legend(title="countries"))+
        labs(title = "Total catch")
    }
    else {
      data <- data %>%
        mutate(perc_disc = (discards/(landings+discards))*100)
      data1 <- data %>%
        group_by(country) %>%
        summarise(avg=mean(perc_disc)) %>%
        arrange(., desc(avg)) %>% 
        top_n(., n=number)
      # filter(avg>30)
      
      data_high <- data %>% filter(country %in% data1$country)
      
      ggplot(data=data_high, aes(x=years, y=perc_disc, colour=country))+
        geom_line(size=1.3)+
        theme(legend.position = "right")+
        labs(title = "Share of discards in total catch", y = "percentage")
      
      # ggplot(data=data_high, aes(x=years, y=perc_disc))+
      #   geom_area(aes(fill=country))+
      #   labs(title = "Discards > 30%",  y="percentage")+
      #   theme(legend.position = "right")
    }
  })
  
  # Output 2: Table ----
  output$values <- renderTable({
    calcTable()
  })
  
}


# Create Shiny app ----
shinyApp(ui = ui, server = server)


```

