---
title: "Sea Around Us Project"
author: "Leonie Grau, Felix Rentschler, Victoria Kolodziej"
date: "March 6, 2018"
output: html_document
runtime: shiny
---

# Introduction

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
The Sea Around Us initiative is a research project by the University of British Columbia, that offers reconstructed catch data for worldwide fisheries and fishery resources. Since overfishing is the major driver on the declining viability of fish stocks, a monitoring program with valid data is crucial for evaluating the marine development. The database provides information for policy makers, fishery managers and the scientific community to get a better understanding of global fishery catches and more detailed information about which countries are involved.  The impact of fishery on the marine ecosystems therefore is presented on a spatial scale in such a way that it has an ecological and political relevance. <cite>(Pauly D. and Zeller D. , 2015. Sea Around Us Concepts, Design and Data<cite> [www.seaaroundus.org](www.seaaroundus.org))

The database covers catch data over 64 years (1950-2014), 4 fishing sectors (industrial, artisanal, subsistence and recreational), 2 catch types (landed and discarded catch) and 2 types of reporting status (reported and unreported). There are 5 regions that can be shown: Exclusive-economic zones (EEZ), large marine ecosystems (LME), regional fisheries management organizations (RFMO),  EEZ-bordering and taxon. [...]

For the purpose of a more in depth data work the R library “seaaroundus” is provided. The aim of this project is to create an interactive web app to present specific data from the sea around us database. This is realised by the use of  the R package “shiny”. 


## Material and Methods

You can embed Shiny inputs and outputs in your document. Outputs are automatically updated whenever inputs change.  This demonstrates how a standard R plot can be made interactive by wrapping it in the Shiny `renderPlot` function. The `selectInput` and `sliderInput` functions create the input widgets used to drive the plot.



## Results

It's also possible to embed an entire Shiny application within an R Markdown document using the `shinyAppDir` function. This example embeds a Shiny application located in another directory:

```{r tabsets, echo=FALSE}

shinyAppDir("/Users/Fex/Studium/UmWi/1. WS 2017:18/6_GeodataVisual/gdbv/",
  options=list(
    width="100%", height=550
  )
) 

```

Note the use of the `height` parameter to determine how much vertical space the embedded application should occupy.

You can also use the `shinyApp` function to define an application inline rather then in an external directory.

In all of R code chunks above the `echo = FALSE` attribute is used. This is to prevent the R code within the chunk from rendering in the document alongside the Shiny components.



